---
title: "IS6481 | Lab 5 -- Unsupervised Learning Responses"
output: 
  html_document: 
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# options(width = 120)
library(ggplot2)
library(fpc)
library(tidyverse)
library(ggrepel)

mm <- src_mysql(
  host='is6481-classroom-data.crximyhxko7e.us-east-1.rds.amazonaws.com',
  port=3306,
  user='admin',
  password='test1234',
  dbname='unsupervised_learning'
)
```

```{r data_acquisition,message=FALSE,warning=FALSE,echo=FALSE}
d_protein <- tbl(mm,'protein') %>% collect() # Get data from the database
```

## To Turn In
This document serves as the template for this assignment. Go through each set of questions and answer them to the best of your ability. Please answer in-line with the bulleted questions. When you are finished, please click the "Knit" button above to preview the final document (should be an HTML document). When finished, download the file from R Studio Cloud by clicking "lab5_turnin.html", click the gear menu labeled "More" and click export. Save the file locally and then upload to Canvas.

### Understanding Density Plots
Answer the following questions:

* *Look at the density plot for white meat, what does it mean? In other words, what does it tell you about white meat consumption?*
* *What would a bad set of density plots look like for a cluster solution?*
* *Do you think this set of variables will provide an interesting solution? Why?*

```{r explore_visually,fig.height=8,fig.width=10,echo=FALSE}
density_data <- d_protein %>% #select(Country,RedMeat,WhiteMeat,Eggs,Milk) %>% 
    pivot_longer(cols=-Country,names_to='type')
ggplot(density_data, aes(x=value)) + 
    geom_density(fill='#F6FEAA',alpha=0.6) + 
    facet_wrap(~type,scales='free') +
    labs(x='Value',y='Density Fuction') +
    theme_minimal() +
    theme(panel.grid.minor.y=element_blank())
```


### Variable Choice and Transformation
As discussed in the lab, we could narrow down our set of variables by combining the two meat categories by taking the row-wise maximum between RedMeat and WhiteMeat.

* *What do you think it would do to the solution to collapse the two meat variables?*

### Cluster Algorithms
As discussed in the lab, there are several different versions of clustering algorithms available to the analyst. They all work differently. The following code runs the k-means cluster algorithm as seen in the lab writeup. We then run a heirarchical cluster algorithm to compare.

```{r scaling_variables,echo=FALSE}
var_list <- c("RedMeat", "WhiteMeat", "Eggs", "Milk", "Fish", "Cereals", "Starch", "Nuts", "FrAndVeg") 

m_protein <- as.data.frame(scale(d_protein[, var_list])) # filters to only the attributes we need and scales them
names(m_protein) <- paste0('st_',var_list)

all_protein <- cbind(d_protein,m_protein)
```

```{r cluster}
set.seed(42)
cluster_model <- kmeans(m_protein, centers=5)
cluster_model$centers %>% t() %>% knitr::kable(caption='k-means Cluster Solution')
```

Give a descriptive name to each of the clusters.

* *Cluster 1*
* *Cluster 2*
* *Cluster 3*
* *Cluster 4*
* *Cluster 5*

Following is a cluster analysis of the same data but using a different clustering algorithm. Look at the profiling table and answer the questions below.
```{r hier_cluster}
d <- dist(m_protein,method='euclidean')
hc1 <- hclust(d,method='ward.D')
h_profiling <- m_protein
h_profiling$hclust <- cutree(hc1,k=5)

h_profile_table <- h_profiling %>% pivot_longer(cols=-hclust) %>%
  pivot_wider(names_from=name,values_from=value,values_fn=list(value=mean))

h_profile_table %>% knitr::kable(caption='Heirarchical Cluster Solution')

```


Give a descriptive name to each of the clusters.

* *Cluster 1*
* *Cluster 2*
* *Cluster 3*
* *Cluster 4*
* *Cluster 5*

*Describe the differences between the two solutions.*


### Change the number of clusters

Copy either of the clustering code segments from above, change the number of clusters and do the following^[Hint: it will likely be easier if you choose a smaller number of clusters].

* *Name each of the clusters*
* *Describe the differences between the two solutions*

